Model_additionalAlts

### 2015-12-01 models/Model_additionalAlts.Rmd
prelim code for simplified gradient descent for alternative dist tuning

```{r}
# liked_loved
combos = data.frame(c(0, 0, 1, 0, 0),
                    c(0, 0.05, 1, .05, 0),
                    c(0, 0.1, 1, .1, 0),
                    c(0, 0.1, 0.9, .1, 0),
                    c(0, 0.2, 0.8, .2, 0),
                    c(0, 0.25, 0.8, .25, 0),
                    c(0.05, 0.1, 1, .1, 0.9),
                    c(0.1, 0.1, 1, .1, 0.1))  
focus_alts("liked_loved", combos=combos, alpha=4, normalize=T)
focus_alts("liked_loved", combos=combos, alpha=8, normalize=F)

@@

>>>>

```{r}

# liked_loved
combos = data.frame(c(0, 0, 1, 0, 0),
                    c(0, 0.05, 1, .05, 0),
                    c(0, 0.1, 1, .1, 0),
                    c(0, 0.1, 0.9, .1, 0),
                    c(0, 0.2, 0.8, .2, 0),
                    c(0, 0.25, 0.8, .25, 0),
                    c(0.05, 0.1, 1, .1, 0.9),
                    c(0.1, 0.1, 1, .1, 0.1))  
focus_alts("liked_loved", combos=combos, alpha=4, normalize=T)
focus_alts("liked_loved", combos=combos, alpha=8, normalize=F)

@@

Fit params - weight update fn()
```{r}
# fit_distr()
# -----------
# target        --> specific scalar family we're interested in (i.e. "some_all")
# numIters      --> 
# eta           --> 
# normalize     --> 
fit_distr = function(target, numIters = 100, eta=0.01,  normalize=F, alpha=1, verbose=F) {
  # Initialize to uniform
  w = rep(0.1, 5)
  # Focus on scalar 'target'
  scalar.focused = data.full.extras[data.full.extras$scale == target, ]
  # Initialize 'mid' alternative

  match.scalars = which(scalar.focused$degree == "hi1" | scalar.focused$degree == "hi2")

  # For numIters iterations...
  for (i in 1:numIters) {
    scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = w
    m = scalar.focused %>%
      do(run.full(., alpha = alpha, addMid=T, normalize=normalize))
    w[1] = w[1] - eta * (sum(m[m$stars == 1, ]$listener.p - m[m$stars == 1, ]$pred))
    w[2] = w[2] - eta * (sum(m[m$stars == 2, ]$listener.p - m[m$stars == 2, ]$pred))
    w[3] = w[3] - eta * (sum(m[m$stars == 3, ]$listener.p - m[m$stars == 3, ]$pred))
    w[4] = w[4] - eta * (sum(m[m$stars == 4, ]$listener.p - m[m$stars == 4, ]$pred))
    w[5] = w[5] - eta * (sum(m[m$stars == 5, ]$listener.p - m[m$stars == 5, ]$pred))
    w = ifelse(w <= 0, 0, w)
    w = ifelse(w >= 1, 1, w)
    if(verbose) {print(w)}
  }
  return(w)
}
normalize = function(w) { return (w / sum(w))}

```

```{r}
targets = c("hi1", "hi2")
fitted.w = fit_distr("good_excellent", numIters=100, alpha=4, eta=0.001, normalize=T)
scalar.focused = data.full.extras[data.full.extras$scale == "good_excellent", ]
scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = fitted.w
m = scalar.focused %>%
  do(run.full(., alpha = 4, addMid=T, normalize=T))
cor(m[m$degree %in% targets, ]$listener.p, m[m$degree %in% targets, ]$pred)
normalize(fitted.w)

fitted.w = fit_distr("liked_loved", numIters=100, alpha=4, eta=0.001, normalize=T)
# fitted.w = fitted.w / sum(fitted.w)
scalar.focused = data.full.extras[data.full.extras$scale == "good_excellent", ]
scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = fitted.w
m = scalar.focused %>%
  do(run.full(., alpha = 4, addMid=T, normalize=T))
cor(m[m$degree %in% targets, ]$listener.p, m[m$degree %in% targets, ]$pred)
normalize(fitted.w)

targets = c("hi1", "low1")
fitted.w = fit_distr("some_all", numIters=200, alpha=4, eta=0.001, normalize=T)
# fitted.w = fitted.w / sum(fitted.w)
scalar.focused = data.full.extras[data.full.extras$scale == "some_all", ]
scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = fitted.w
m = scalar.focused %>%
  do(run.full(., alpha = 4, addMid=T, normalize=T))
cor(m[m$degree %in% targets, ]$listener.p, m[m$degree %in% targets, ]$pred)
normalize(fitted.w)
qplot(stars, listener.p, col=degree, 
      data=m[m$degree %in% targets,],
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

targets = c("hi1", "hi2")
fitted.w = fit_distr("palatable_delicious", numIters=200, alpha=4, eta=0.001, normalize=T)
# fitted.w = fitted.w / sum(fitted.w)
scalar.focused = data.full.extras[data.full.extras$scale == "palatable_delicious", ]
scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = fitted.w
m = scalar.focused %>%
  do(run.full(., alpha = 4, addMid=T, normalize=T))
cor(m[m$degree %in% targets, ]$listener.p, m[m$degree %in% targets, ]$pred)
normalize(fitted.w)

qplot(stars, listener.p, col=degree, 
      data=m[m$degree %in% targets,],
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

targets = c("hi1", "hi2")
fitted.w = fit_distr("memorable_unforgettable", numIters=200, alpha=4, eta=0.01, normalize=T)
# fitted.w = fitted.w / sum(fitted.w)
scalar.focused = data.full.extras[data.full.extras$scale == "memorable_unforgettable", ]
scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = fitted.w
m = scalar.focused %>%
  do(run.full(., alpha = 4, addMid=T, normalize=T))
cor(m[m$degree %in% targets, ]$listener.p, m[m$degree %in% targets, ]$pred)

normalize(fitted.w)

qplot(stars, listener.p, col=degree, 
      data=m[m$degree %in% targets,],
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```
